---
title: "AEGIS Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
library(Seurat)
```

## Purpose

AEGIS is an MVP package for auditing and comparing spatial deconvolution outputs on Seurat spatial objects.

## Object structure

An `aegis` object is a lightweight S3 list:

- `seu`: Seurat spatial object
- `deconv`: named list of method result matrices
- `markers`: marker list
- `audit`: audit outputs (`basic`, `marker`, `spatial`)
- `consensus`: comparison and consensus outputs
- `meta`: run metadata

## Quick workflow on built-in example data

```{r}
data("aegis_example", package = "AEGIS")
markers <- readRDS(system.file("extdata", "marker_list.rds", package = "AEGIS"))

deconv <- simulate_deconv_results(aegis_example, seed = 123)
obj <- as_aegis(aegis_example, deconv, markers = markers)

obj <- audit_basic(obj)
obj <- audit_marker(obj)
obj <- audit_spatial(obj)
obj <- compare_methods(obj)
obj <- compute_consensus(obj)
```

## Audit workflow

```{r, fig.width=7, fig.height=6}
Seurat::SpatialFeaturePlot(aegis_example, features = "nCount_Spatial")
```

```{r, fig.width=8, fig.height=5}
plot_audit(obj, "dominance")
```

```{r, fig.width=8, fig.height=5}
plot_audit(obj, "marker")
```

## Comparison workflow

```{r, fig.width=6, fig.height=5}
plot_compare(obj, "heatmap")
```

## Report generation

```{r eval=FALSE}
render_report(obj, output_file = "aegis_report.html")
```

The report is based on `inst/templates/aegis_report.Rmd` and renders in headless environments.

## Tutorial and citation

- Full tutorial: `vignettes/AEGIS-complete-tutorial.Rmd`
- Repository: <https://github.com/JamesWu7/AEGIS>
- Citation:

```{r eval=FALSE}
citation("AEGIS")
```
