---
title: "AEGIS Complete Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Complete Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
```

## 1. Load Example Seurat Object

```{r}
data("aegis_example", package = "AEGIS")
seu <- aegis_example
```

## 2. Simulated Workflow (Development Path)

```{r}
markers <- readRDS(system.file("extdata", "marker_list.rds", package = "AEGIS"))
deconv_sim <- simulate_deconv_results(
  seu,
  methods = c("RCTD", "SPOTlight", "cell2location"),
  seed = 2026
)

obj_sim <- run_aegis(seu, deconv = deconv_sim, markers = markers)
```

```{r}
knitr::kable(obj_sim$audit$basic$summary)
```

```{r, fig.width=7, fig.height=5}
plot_audit(obj_sim, type = "entropy", method = "SPOTlight")
```

## 3. Real Import Workflow (P8 Adapter Path)

This section mimics exported backend result files and imports them using AEGIS adapter readers.

```{r}
spots <- colnames(seu)[1:8]
seu_small <- suppressWarnings(seu[, spots])
```

### 3.1 Build Example Exported Files

```{r}
tmp_rctd <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    barcode = spots,
    B_cell = c(0.5, 0.2, 0.4, 0.3, 0.1, 0.2, 0.4, 0.5),
    T_cell = c(0.3, 0.6, 0.4, 0.4, 0.7, 0.6, 0.3, 0.3),
    Myeloid = c(0.2, 0.2, 0.2, 0.3, 0.2, 0.2, 0.3, 0.2),
    check.names = FALSE
  ),
  tmp_rctd,
  row.names = FALSE
)

tmp_spotlight <- tempfile(fileext = ".tsv")
utils::write.table(
  data.frame(
    spot_id = spots,
    B_cell = c(0.4, 0.3, 0.5, 0.3, 0.2, 0.3, 0.4, 0.4),
    T_cell = c(0.4, 0.5, 0.3, 0.4, 0.6, 0.5, 0.4, 0.3),
    Myeloid = c(0.2, 0.2, 0.2, 0.3, 0.2, 0.2, 0.2, 0.3),
    sample = "S1",
    check.names = FALSE
  ),
  tmp_spotlight,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE
)

tmp_cell2location <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    spot = spots,
    B_cell = c(15, 8, 12, 10, 5, 7, 9, 11),
    T_cell = c(12, 18, 10, 13, 20, 16, 12, 10),
    Myeloid = c(5, 6, 4, 8, 5, 7, 6, 9),
    x = seq_along(spots),
    y = rev(seq_along(spots)),
    check.names = FALSE
  ),
  tmp_cell2location,
  row.names = FALSE
)
```

### 3.2 Import and Standardize

```{r}
rctd <- read_rctd(tmp_rctd)
spotlight <- read_spotlight(tmp_spotlight)
cell2location <- read_cell2location(tmp_cell2location)
```

### 3.3 Build AEGIS Object from Real Inputs

```{r}
obj_real <- run_aegis(
  seu_small,
  deconv = list(
    RCTD = rctd,
    SPOTlight = spotlight,
    cell2location = cell2location
  ),
  do_marker = FALSE,
  do_spatial = FALSE
)
```

```{r}
knitr::kable(obj_real$audit$basic$summary)
```

### 3.4 Additional Import Adapters (same standardized output shape)

```{r}
tmp_card <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    spot_id = spots,
    B_cell = c(0.55, 0.25, 0.45, 0.35, 0.15, 0.25, 0.35, 0.45),
    T_cell = c(0.30, 0.55, 0.35, 0.40, 0.65, 0.55, 0.45, 0.35),
    Myeloid = c(0.15, 0.20, 0.20, 0.25, 0.20, 0.20, 0.20, 0.20),
    check.names = FALSE
  ),
  tmp_card,
  row.names = FALSE
)

tmp_destvi <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    barcode = spots,
    B_cell = c(20, 8, 12, 9, 5, 7, 8, 10),
    T_cell = c(10, 18, 11, 13, 20, 17, 12, 9),
    Myeloid = c(4, 6, 5, 7, 4, 6, 5, 8),
    check.names = FALSE
  ),
  tmp_destvi,
  row.names = FALSE
)

tmp_stdec <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    spot = spots,
    topic1 = c(0.6, 0.3, 0.4, 0.5, 0.2, 0.3, 0.4, 0.5),
    topic2 = c(0.4, 0.7, 0.6, 0.5, 0.8, 0.7, 0.6, 0.5),
    check.names = FALSE
  ),
  tmp_stdec,
  row.names = FALSE
)

card <- read_card(tmp_card)
destvi <- read_destvi(tmp_destvi)
stdec <- read_stdeconvolve(tmp_stdec)

obj_extra <- as_aegis(
  seu_small,
  deconv = list(CARD = card, DestVI = destvi, STdeconvolve = stdec)
)
obj_extra <- audit_basic(obj_extra)
knitr::kable(obj_extra$audit$basic$summary)
```

## 4. Multi-sample Workflow (P6)

```{r}
spots_all <- colnames(seu)
n_half <- floor(length(spots_all) / 2)

seu_list <- list(
  sample1 = suppressWarnings(seu[, spots_all[seq_len(n_half)]]),
  sample2 = suppressWarnings(seu[, spots_all[seq.int(n_half + 1L, length(spots_all))]])
)

deconv_nested <- list(
  sample1 = simulate_deconv_results(seu_list$sample1, methods = c("RCTD", "SPOTlight"), seed = 333),
  sample2 = simulate_deconv_results(seu_list$sample2, methods = c("RCTD", "SPOTlight"), seed = 444)
)

obj_multi <- run_aegis(
  seu_list,
  deconv = deconv_nested,
  markers = markers
)

sample_summary <- summarize_by_sample(obj_multi)
knitr::kable(sample_summary)
```

```{r eval=FALSE}
render_report_batch(obj_multi, output_dir = "reports")
```

## 5. Optional Report Generation

```{r eval=FALSE}
render_report(obj_sim, output_file = "aegis_report.html")
```

## 6. Summary

- Use `simulate_deconv_results()` for reproducible demos and method development.
- Use method adapters (`read_rctd()`, `read_spotlight()`, `read_cell2location()`, `read_card()`, `read_destvi()`, `read_stdeconvolve()`, etc.) for external exports.
- Use `run_aegis()` as the unified pipeline entry for single-sample and multi-sample analysis.
- Use `summarize_by_sample()` and `render_report_batch()` for multi-sample projects.
