---
title: "AEGIS One-Step Deconvolution Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS One-Step Deconvolution Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
```

## 1) What one-step deconvolution means in AEGIS

AEGIS supports three method support modes:
- `run_and_import_r`: runnable directly in R
- `run_and_import_python`: runnable through Python (optional)
- `import_only`: import exported outputs only

Use the registry first:

```{r}
registry <- get_supported_methods()
knitr::kable(registry)
```

## 2) Load spatial data and markers

```{r}
data("aegis_example", package = "AEGIS")
seu <- aegis_example
markers <- readRDS(system.file("extdata", "marker_list.rds", package = "AEGIS"))
```

## 3) Prepare a reference object

In practice this should be your real single-cell reference object.

```{r}
reference <- matrix(
  c(4, 2, 1, 2, 5, 2, 1, 2, 6),
  nrow = 3,
  byrow = TRUE
)
rownames(reference) <- c("B_cell", "T_cell", "Myeloid")
colnames(reference) <- c("g1", "g2", "g3")
```

## 4) Run one method (direct execution)

```{r eval=FALSE}
res_one <- run_deconvolution(
  seu = seu,
  reference = reference,
  methods = "SPOTlight",
  strict = TRUE
)

obj_one <- run_aegis(res_one$seu, deconv = res_one$deconv, markers = markers)
```

## 5) Run multiple methods in one step

```{r eval=FALSE}
res_multi <- run_deconvolution(
  seu = seu,
  reference = reference,
  methods = c("SPOTlight", "CARD", "RCTD"),
  strict = FALSE
)

obj_multi <- run_aegis(res_multi$seu, deconv = res_multi$deconv, markers = markers)
```

## 6) Full one-step wrapper

```{r eval=FALSE}
obj_full <- run_aegis_full(
  seu = seu,
  reference = reference,
  methods = c("SPOTlight", "CARD", "RCTD"),
  markers = markers,
  strict = FALSE
)
```

## 7) Optional Python-backed methods

Python-backed methods are optional. If Python modules are missing:
- strict mode: clear error
- non-strict mode: method is skipped with a message

```{r eval=FALSE}
res_py <- run_deconvolution(
  seu = seu,
  reference = reference,
  methods = c("cell2location", "DestVI", "Tangram"),
  use_python = TRUE,
  strict = FALSE
)
```

## 8) Downstream AEGIS analysis and visualization

For a fully reproducible vignette run (without backend dependencies), this chunk
uses simulated deconvolution and then runs the same downstream pipeline.

```{r}
demo_methods <- registry$method_name
deconv_demo <- simulate_deconv_results(
  seu,
  methods = demo_methods,
  cell_types = c("B_cell", "T_cell", "Myeloid"),
  seed = 909
)

obj <- run_aegis(seu, deconv = deconv_demo, markers = markers)
obj <- score_methods(obj)
obj <- rank_methods(obj, method = "mean_rank")
obj <- compute_consensus(obj, strategy = "weighted", top_n = min(4, length(demo_methods)))

obj$consensus$result$methods_used
```

```{r, fig.width=14, fig.height=6.5}
plot_compare(obj, type = "heatmap")
```

```{r, fig.width=7.2, fig.height=5.6}
plot_compare(obj, type = "consensus_map")
```

```{r, fig.width=9, fig.height=6.2}
plot_compare(obj, type = "ranking")
```

```{r, fig.width=7.2, fig.height=5.6}
plot_compare(obj, type = "disagreement_map")
```

```{r, fig.width=7.2, fig.height=5.6}
plot_compare(obj, type = "confidence_map")
```

## 9) Practical notes

1. Use direct execution when method dependencies and inputs are ready.
2. Use importer functions (`read_*`) when methods are not executable in your environment.
3. Treat Python-backed methods as optional unless you manage a validated Python environment.
4. Use `run_aegis_full()` for one-step execution plus downstream AEGIS analysis.
