---
title: "AEGIS Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
```

## Purpose

AEGIS is designed to audit and compare spatial deconvolution outputs on Seurat spatial objects.

Current MVP workflows:

1. Simulate deconvolution outputs for demos and method development.
2. Import real exported results from RCTD/SPOTlight/cell2location.
3. Run a unified audit/comparison/consensus pipeline on the standardized matrices.

## AEGIS Object Structure

`as_aegis()` creates a lightweight S3 object:

- `seu`: Seurat object
- `deconv`: named list of spot-by-celltype matrices
- `markers`: optional marker list
- `audit`: audit results (`basic`, `marker`, `spatial`)
- `consensus`: comparison + consensus outputs
- `meta`: user metadata list

## Simulated Workflow

```{r}
data("aegis_example", package = "AEGIS")

deconv <- simulate_deconv_results(aegis_example, seed = 123)
obj <- as_aegis(aegis_example, deconv, markers = aegis_default_markers())
obj <- audit_basic(obj)
obj <- audit_marker(obj)
obj <- audit_spatial(obj)
obj <- compare_methods(obj)
obj <- compute_consensus(obj)

knitr::kable(obj$audit$basic$summary)
```

## Real Import Workflow

AEGIS imports exported tables and does not run external deconvolution backends.

```{r}
spots <- colnames(aegis_example)[1:6]

tmp_rctd <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    barcode = spots,
    B_cell = c(0.5, 0.3, 0.2, 0.6, 0.1, 0.4),
    T_cell = c(0.3, 0.4, 0.4, 0.2, 0.7, 0.4),
    Myeloid = c(0.2, 0.3, 0.4, 0.2, 0.2, 0.2),
    check.names = FALSE
  ),
  tmp_rctd,
  row.names = FALSE
)

rctd <- read_rctd(tmp_rctd)
obj_real <- as_aegis(aegis_example[, spots], deconv = list(RCTD = rctd))
obj_real <- audit_basic(obj_real)
knitr::kable(obj_real$audit$basic$summary)
```

## Next Steps

See the demo and complete tutorials for full end-to-end examples and multi-method import workflows.
