---
title: "AEGIS Complete Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Complete Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
library(Seurat)
```

## 1. Purpose

This tutorial is a complete, end-to-end AEGIS workflow for spatial transcriptomics deconvolution auditing.

## 2. Inputs

- Primary input is a Seurat spatial object.
- Built-in example object: `aegis_example`.
- Marker list: `inst/extdata/marker_list.rds`.

```{r}
data("aegis_example", package = "AEGIS")
markers <- readRDS(system.file("extdata", "marker_list.rds", package = "AEGIS"))
```

## 3. Slice visualization (spatial group section)

```{r, fig.width=7, fig.height=6}
Seurat::SpatialFeaturePlot(aegis_example, features = "nCount_Spatial")
```

## 4. Simulate methods and create AEGIS object

```{r}
deconv <- simulate_deconv_results(
  aegis_example,
  methods = c("RCTD", "SPOTlight", "cell2location"),
  seed = 2026
)

obj <- as_aegis(aegis_example, deconv, markers = markers)
obj
```

## 5. Run quality audits

```{r}
obj <- audit_basic(obj)
obj <- audit_marker(obj)
obj <- audit_spatial(obj)
```

```{r, fig.width=8, fig.height=5}
plot_audit(obj, "dominance")
```

```{r, fig.width=8, fig.height=5}
plot_audit(obj, "marker")
```

## 6. Compare methods and compute consensus

```{r}
obj <- compare_methods(obj)
obj <- compute_consensus(obj)
```

```{r, fig.width=6, fig.height=5}
plot_compare(obj, "heatmap")
```

```{r, fig.width=7, fig.height=5}
plot_compare(obj, "consensus_map")
```

## 7. Render standalone report

```{r eval=FALSE}
render_report(obj, output_file = "aegis_tutorial_report.html")
```

## 8. Citation and repository

```{r eval=FALSE}
citation("AEGIS")
```

Repository: <https://github.com/JamesWu7/AEGIS>
