---
title: "AEGIS Demo Human Lymph Node"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Demo Human Lymph Node}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
library(Seurat)
```

## Demo data source

This vignette uses the built-in Human Lymph Node example object (`aegis_example`) so it can render in CI, pkgdown, and source builds.

```{r}
data("aegis_example", package = "AEGIS")
seu <- aegis_example
seu
```

## Optional: load directly from authoritative raw files

If you are in the repository root and have these raw files available:

- `V1_Human_Lymph_Node_filtered_feature_bc_matrix.h5`
- `V1_Human_Lymph_Node_spatial.tar.gz`
- `V1_Human_Lymph_Node_metrics_summary.csv`

then you can load from disk with:

```{r eval=FALSE}
seu <- load_10x_lymphnode(data_dir = ".")
```

## Spatial transcriptomics slice example

```{r, fig.width=7, fig.height=6}
Seurat::SpatialFeaturePlot(seu, features = "nCount_Spatial")
```

## Simulate deconvolution and run AEGIS

```{r}
deconv <- simulate_deconv_results(seu, seed = 2026)
obj <- as_aegis(seu, deconv)
obj <- audit_basic(obj)
```

## Basic audit output

```{r}
knitr::kable(obj$audit$basic$summary)
```
