---
title: "AEGIS Demo: Human Lymph Node"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Demo Human Lymph Node}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
library(Seurat)
```

## Demo Data Source

This tutorial uses the built-in Human Lymph Node example object (`aegis_example`) so it can run in CI and pkgdown builds.

```{r}
data("aegis_example", package = "AEGIS")
seu <- aegis_example
seu
```

## Optional: Load from Authoritative Raw Files

If you are in repository root and have the authoritative raw files:

- `V1_Human_Lymph_Node_filtered_feature_bc_matrix.h5`
- `V1_Human_Lymph_Node_spatial.tar.gz`
- `V1_Human_Lymph_Node_metrics_summary.csv`

you can load directly with:

```{r eval=FALSE}
seu <- load_10x_lymphnode(data_dir = ".")
```

## Tissue Context

```{r, fig.width=7, fig.height=6}
Seurat::SpatialFeaturePlot(seu, features = "nCount_Spatial")
```

## End-to-End AEGIS Workflow (Simulated Inputs)

```{r}
markers <- readRDS(system.file("extdata", "marker_list.rds", package = "AEGIS"))
deconv <- simulate_deconv_results(seu, seed = 2026)

obj <- as_aegis(seu, deconv, markers = markers)
obj <- audit_basic(obj)
obj <- audit_marker(obj)
obj <- audit_spatial(obj)
obj <- compare_methods(obj)
obj <- compute_consensus(obj)
```

```{r}
knitr::kable(obj$audit$basic$summary)
```

## Example Plots

```{r, fig.width=7, fig.height=5}
plot_audit(obj, type = "dominance", method = "RCTD")
```

```{r, fig.width=7, fig.height=5}
plot_compare(obj, type = "heatmap")
```

## Notes

- AEGIS can run with simulated outputs or imported real exported tables.
- AEGIS does not execute external deconvolution methods.
