---
title: "AEGIS Demo: Human Lymph Node"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Demo Human Lymph Node}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
library(Seurat)

safe_tissue_context_plot <- function(seu) {
  has_spatial_image <- FALSE
  has_ncount <- "nCount_Spatial" %in% colnames(seu[[]])

  has_spatial_image <- tryCatch(
    length(Seurat::Images(seu)) > 0L,
    error = function(e) FALSE
  )

  if (has_spatial_image && has_ncount) {
    return(Seurat::SpatialFeaturePlot(seu, features = "nCount_Spatial"))
  }

  if (has_ncount) {
    return(
      ggplot2::ggplot(seu[[]], ggplot2::aes(x = .data$nCount_Spatial)) +
        ggplot2::geom_histogram(bins = 30, fill = "#1f6fb2", color = "white") +
        ggplot2::theme_minimal(base_size = 12) +
        ggplot2::labs(
          title = "Fallback tissue context view",
          subtitle = "Spatial image not available; showing nCount_Spatial distribution",
          x = "nCount_Spatial",
          y = "Spots"
        )
    )
  }

  ggplot2::ggplot(data.frame(spot = seq_len(ncol(seu))), ggplot2::aes(x = spot, y = 1)) +
    ggplot2::geom_point(size = 1.3, color = "#1f6fb2") +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::labs(
      title = "Fallback tissue context view",
      subtitle = "No spatial image and no nCount_Spatial metadata; plotting spot index",
      x = "Spot index",
      y = NULL
    )
}
```

## Demo Data Source

This tutorial uses the built-in Human Lymph Node example object (`aegis_example`) so it can run in CI and pkgdown builds.

```{r}
data("aegis_example", package = "AEGIS")
seu <- aegis_example
seu
```

## Optional: Load from Authoritative Raw Files

If you are in repository root and have the authoritative raw files:

- `V1_Human_Lymph_Node_filtered_feature_bc_matrix.h5`
- `V1_Human_Lymph_Node_spatial.tar.gz`
- `V1_Human_Lymph_Node_metrics_summary.csv`

you can load directly with:

```{r eval=FALSE}
seu <- load_10x_lymphnode(data_dir = ".")
```

## Tissue Context

```{r, fig.width=7, fig.height=6}
safe_tissue_context_plot(seu)
```

## End-to-End AEGIS Workflow (Simulated Inputs)

```{r}
markers <- readRDS(system.file("extdata", "marker_list.rds", package = "AEGIS"))
deconv <- simulate_deconv_results(seu, seed = 2026)

obj <- run_aegis(seu, deconv = deconv, markers = markers)
```

```{r}
knitr::kable(obj$audit$basic$summary)
```

## Example Plots

```{r, fig.width=7, fig.height=5}
plot_audit(obj, type = "dominance", method = "RCTD")
```

```{r, fig.width=7, fig.height=5}
plot_compare(obj, type = "heatmap")
```

## Multi-sample Example (Two Sections from the Demo Object)

```{r}
spots_all <- colnames(seu)
n_half <- floor(length(spots_all) / 2)

seu_list <- list(
  section_A = suppressWarnings(seu[, spots_all[seq_len(n_half)]]),
  section_B = suppressWarnings(seu[, spots_all[seq.int(n_half + 1L, length(spots_all))]])
)

deconv_nested <- list(
  section_A = simulate_deconv_results(seu_list$section_A, methods = c("RCTD", "SPOTlight"), seed = 91),
  section_B = simulate_deconv_results(seu_list$section_B, methods = c("RCTD", "SPOTlight"), seed = 92)
)

obj_multi <- run_aegis(seu_list, deconv = deconv_nested, markers = markers)
knitr::kable(summarize_by_sample(obj_multi))
```

## Notes

- AEGIS can run with simulated outputs or imported real exported tables.
- AEGIS does not execute external deconvolution methods.
- Multi-sample projects can use `run_aegis()` + `summarize_by_sample()` + `render_report_batch()`.
