---
title: "AEGIS Complete Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEGIS Complete Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(AEGIS)
```

## 1. Load Example Seurat Object

```{r}
data("aegis_example", package = "AEGIS")
seu <- aegis_example
```

## 2. Simulated Workflow (Development Path)

```{r}
markers <- readRDS(system.file("extdata", "marker_list.rds", package = "AEGIS"))
deconv_sim <- simulate_deconv_results(
  seu,
  methods = c("RCTD", "SPOTlight", "cell2location"),
  seed = 2026
)

obj_sim <- as_aegis(seu, deconv_sim, markers = markers)
obj_sim <- audit_basic(obj_sim)
obj_sim <- audit_marker(obj_sim)
obj_sim <- audit_spatial(obj_sim)
obj_sim <- compare_methods(obj_sim)
obj_sim <- compute_consensus(obj_sim)
```

```{r}
knitr::kable(obj_sim$audit$basic$summary)
```

```{r, fig.width=7, fig.height=5}
plot_audit(obj_sim, type = "entropy", method = "SPOTlight")
```

## 3. Real Import Workflow (P5 Path)

This section mimics exported backend result files and imports them using the new readers.

```{r}
spots <- colnames(seu)[1:8]
seu_small <- suppressWarnings(seu[, spots])
```

### 3.1 Build Example Exported Files

```{r}
tmp_rctd <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    barcode = spots,
    B_cell = c(0.5, 0.2, 0.4, 0.3, 0.1, 0.2, 0.4, 0.5),
    T_cell = c(0.3, 0.6, 0.4, 0.4, 0.7, 0.6, 0.3, 0.3),
    Myeloid = c(0.2, 0.2, 0.2, 0.3, 0.2, 0.2, 0.3, 0.2),
    check.names = FALSE
  ),
  tmp_rctd,
  row.names = FALSE
)

tmp_spotlight <- tempfile(fileext = ".tsv")
utils::write.table(
  data.frame(
    spot_id = spots,
    B_cell = c(0.4, 0.3, 0.5, 0.3, 0.2, 0.3, 0.4, 0.4),
    T_cell = c(0.4, 0.5, 0.3, 0.4, 0.6, 0.5, 0.4, 0.3),
    Myeloid = c(0.2, 0.2, 0.2, 0.3, 0.2, 0.2, 0.2, 0.3),
    sample = "S1",
    check.names = FALSE
  ),
  tmp_spotlight,
  sep = "\t",
  quote = FALSE,
  row.names = FALSE
)

tmp_cell2location <- tempfile(fileext = ".csv")
utils::write.csv(
  data.frame(
    spot = spots,
    B_cell = c(15, 8, 12, 10, 5, 7, 9, 11),
    T_cell = c(12, 18, 10, 13, 20, 16, 12, 10),
    Myeloid = c(5, 6, 4, 8, 5, 7, 6, 9),
    x = seq_along(spots),
    y = rev(seq_along(spots)),
    check.names = FALSE
  ),
  tmp_cell2location,
  row.names = FALSE
)
```

### 3.2 Import and Standardize

```{r}
rctd <- read_rctd(tmp_rctd)
spotlight <- read_spotlight(tmp_spotlight)
cell2location <- read_cell2location(tmp_cell2location)
```

### 3.3 Build AEGIS Object from Real Inputs

```{r}
obj_real <- as_aegis(
  seu_small,
  deconv = list(
    RCTD = rctd,
    SPOTlight = spotlight,
    cell2location = cell2location
  )
)
obj_real <- audit_basic(obj_real)
obj_real <- compare_methods(obj_real)
obj_real <- compute_consensus(obj_real)
```

```{r}
knitr::kable(obj_real$audit$basic$summary)
```

## 4. Optional Report Generation

```{r eval=FALSE}
render_report(obj_sim, output_file = "aegis_report.html")
```

## 5. Summary

- Use `simulate_deconv_results()` for reproducible demos and method development.
- Use `read_rctd()`, `read_spotlight()`, `read_cell2location()` for external exports.
- Use the same downstream pipeline (`audit_*`, `compare_methods`, `compute_consensus`) in both paths.
